\begin{resumo}[Abstract]
 \begin{otherlanguage*}{english}
   The main objective of this research was the construction of a Security Catalog for the MVC (Model-View-Controller) Architectural Pattern, 
   aiming at assisting Requirements Engineers and Software Engineers in specifying the requirements associated with Security, a non-functional requirement. The catalog is documented based on a graph-oriented modeling of interdependence of non-functional requirements, taking into account good practices of Goal-Oriented Requirements Engineering. In the construction of this catalog, some pillars were considered, such as concepts associated to Security and substantiated on authors of the area, with emphasis on Confidentiality, Integrity and Availability, in addition to the definition of Information Security established by ISO 27001. As an extension of the catalog's base contributions, it has been mapped to/associated with the MVC pattern layers in order to further assist catalog users while developing their applications based on this three-tier Architectural Pattern. Lastly, the catalog has been applied in different usage scenarios, which has made possible not only a catalog with a high level of abstraction, based on the interdependencies of the non-functional requirements correlated with Security, but also a catalog that agrees a series of operations, which they seek to fulfill (at least satisfactory) with the specified non-functional requirements.

   \vspace{\onelineskip}
 
   \noindent 
   \textbf{Key-words}: Software Security Catalog. Architectural Pattern Model-View-Controller. Goal-Oriented Requirements Engineering. Security. Confidentiality. Integrity. Availability. 
 \end{otherlanguage*}
\end{resumo}
